---
title: "Linear Regression Model"
author: "Bowei Zhang"
date: "2023-03-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setenv(LANG = "en")
library(alr4)
library(tidyverse)
library(corrgram)
library(Metrics)
library(GGally)
source("data_preprocessing.R")
```

# Data Wrangling 
```{r}
# See distribution
hist(df$area)
```
```{r}
# Distribution without area = 0.1
hist(df[df$area != 0.1,]$area, breaks = 20)
```
### update 'df'
```{r}
df <- df %>%
  dplyr::select(-area_size) %>%
  filter(area != 0.1) %>%
  mutate(log.area = log(area))
head(df)
hist(df$log.area)
```
# Data Analysis

## variable selection
```{r}
step(lm(log.area ~ 1, data = df %>% dplyr::select(-area)), 
     ~ X + Y + FFMC + DMC + DC + ISI + temp + RH + wind, direction="both", trace = 0)
```
```{r}
m1 = lm(formula = log.area ~ Y + ISI, data = df)
summary(m1)
```
```{r}
par(mfrow = c(2,2))
plot(m1)
```
Adjusted R-square is too low.

## PCA
```{r, echo = FALSE}
# change categorical variables back to continuous variables
df1.cont <- df
df1.cont$X <- as.numeric(df1.cont$X)
df1.cont$Y <- as.numeric(df1.cont$Y)
df1.cont$month <- as.numeric(df1.cont$month)
df1.cont$day <- as.numeric(df1.cont$day)
# PCA including response
# cannot contain categorical variables
full.pca <- prcomp(df1.cont %>% select(-X, -month, -day), center = TRUE, scale. = TRUE) 
summary(full.pca)
round(full.pca$rotation,3)
```
## During Summer
```{r, eval = TRUE, echo = FALSE, out.width="50%"}
# Focus on data in the summer
df_summer <- df %>%
  filter(month %in% c('jun', 'jul', 'aug', 'sep', 'oct'))
plot(log.area ~ month, data = df_summer)
# change the month back to continuous variable
df_summer$month <- as.numeric(df_summer$month)
```
```{r, eval = TRUE, echo= FALSE}
# Polynomial  
m.summer1 <- lm(log.area ~ month, data = df_summer)
summary(m.summer1)
m.summer2 <- lm(log.area ~ poly(month,2), data = df_summer)
summary(m.summer2)
# Diagnostics
par(mfrow = c(2,2))
plot(m.summer1)
```

# Conclusion 
* According to our results, the area burned from forest fires can not be predicted using linear regression model from the variables that we havein the dataset.
* Other variables - such as elevation, levels of human activity - may be necessary to predict area burned
* Other statistical reports have had similar diffuculties predicting area burned.[(source)](https://www.kaggle.com/code/psvishnu/forestfire-impact-prediction-stats-and-ml/notebook)
* Some suggest that neural network and other Tree based models could give a better result.